\vspace{1.0cm} \LARGE \bf MERS-CoV spillover at the camel-human interface

### 2017-08-05 mers-structure.tex
Add N vs Ne text

Despite lack of convergence, neither of the two demographic reconstructions show evidence of fluctuations in the relative genetic diversity ($N_e \tau$) of  MERS-CoV over time (Figure \ref{skygrid}).
However, we do note that estimates of relative genetic diversity are relatively low overall, and MERS-CoV phylogeny resembles a ladder often seen in human influenza A virus phylogenies \citep{bedford_strength_2011}.
This raises the possibility that MERS-CoV populations undergo lineage turnover at a rate not often seen in other livestock viruses (\textit{e.g.} swine influenza A viruses \citep{vijaykrishna_long-term_2011}).
Part of this could be caused by MERS-CoV ability to re-infect camels previously exposed to the virus \citep{ali_systematic_2017}, which may result in a process akin to antigenic drift, where the depletion of susceptible calves leads to ever increasing selective pressure for the virus to evolve antigenically.
While this could be tested by looking for seasonally varying patterns in selection on MERS-CoV surface proteins, recombination would remain a key confounding problem.
\tbc{Include results of N and S to indicate that surface protein is evolving quickly supporting antigenic evolution. Or not, if the data says otherwise...}

>>>>

Despite lack of convergence, neither of the two demographic reconstructions show evidence of fluctuations in the relative genetic diversity ($N_e \tau$) of  MERS-CoV over time (Figure \ref{skygrid}).
However, we do note that estimates of relative genetic diversity are relatively low overall with a mean estimate of $N_e\tau$ of 3.49 years, and consequently MERS-CoV phylogeny resembles a ladder often seen in human influenza A virus phylogenies \citep{bedford_strength_2011}.
This estimate of $N_e\tau$ can be translated into an estimate of effective prevalence $N_e$ given an estimate of generation time or serial intervals.
Given a 20 day duration of infection in camels \tbc{please add citation}, we expect a 10 day serial interval or $\tau$ of 0.027 years per generation.
Dividing by $\tau$ yields an estimate of $N_e$ of 129.3 individuals.

Here, we compare this estimate of effective prevalence to a rough estimate of census prevalence.
Given extremely high seroprevalence estimates within camels in Saudi Arabia \tbc{please add citation}, we expect camels to usually be infected within their first year of life.
Therefore we can estimate the rough number of camel infections per year as the number of calves produced each year.
We find there are 270 000 camels in Saudi Arabia and that in camel farming practices a camel female can be bred for about 22 years and about 10 calves can be produced within the 22 year period \tbc{cite Elmi 1998}.
This gives an expectation of each female camel producing, on average, 0.45 calves per year.
Thus, we expect $270000 \times 0.5 \times 0.45 = 60750$ new calves produced yearly and correspondingly 60750 new infections every year, which spread over 10 day serial intervals gives an average of $N = 1660$ infections.

This raises the possibility that MERS-CoV populations undergo lineage turnover at a rate not often seen in other livestock viruses (\textit{e.g.} swine influenza A viruses \citep{vijaykrishna_long-term_2011}).
Part of this could be caused by MERS-CoV ability to re-infect camels previously exposed to the virus \citep{ali_systematic_2017}, which may result in a process akin to antigenic drift, where the depletion of susceptible calves leads to ever increasing selective pressure for the virus to evolve antigenically.
While this could be tested by looking for seasonally varying patterns in selection on MERS-CoV surface proteins, recombination would remain a key confounding problem.
\tbc{Include results of N and S to indicate that surface protein is evolving quickly supporting antigenic evolution. Or not, if the data says otherwise...}
